### 2013-12-26 tex/ms.tex
streams: mention that we model leading and trailing separately, sign for leading/trailing, some more tweaks

\documentclass[12pt,preprint]{aastex}
%\documentclass{emulateapj}
\usepackage{amssymb,amsmath}
%\usepackage{caption}
%\DeclareCaptionLabelSeparator{dotemdash}{.--- }
%\captionsetup[figure]{labelformat=simple, labelsep=dotemdash}
%\usepackage{color,hyperref}
% hypertex insanity
%\definecolor{linkcolor}{rgb}{0,0,0.25}
%\hypersetup{
%  colorlinks=true,        % false: boxed links; true: colored links
%  linkcolor=linkcolor,    % color of internal links
%  citecolor=linkcolor,    % color of links to bibliography
%  filecolor=linkcolor,    % color of file links
%  urlcolor=linkcolor      % color of external links
%}
\newcounter{address}
\setcounter{address}{1}
%\usepackage{sidecap}
\setlength{\emergencystretch}{2em}%No overflowing references
\newcommand{\ie}{i.e.}
\newcommand{\etal}{et al.}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\eg}{e.g.}
\newcommand{\eqnname}{equation}
\newcommand{\Eqnname}{Equation}
\newcommand{\equationname}{\eqnname}
\renewcommand{\tablename}{Table}
\renewcommand{\figurename}{Figure}
\newcommand{\figurenames}{\figurename~s}
\newcommand{\sectionname}{$\mathsection$}
\newcommand{\normal}{\ensuremath{\mathcal{N}}}
\newcommand{\flag}[1]{\texttt{\lowercase{#1}}}
\newcommand{\feh}{\ensuremath{[\mathrm{Fe/H}]}}
\newcommand{\afe}{\ensuremath{[\alpha\mathrm{/Fe}]}}
\newcommand{\logg}{log g}
\newcommand{\Ro}{\ensuremath{R_0}}

\renewcommand{\vec}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\unitvec}[1]{\ensuremath{\mathbf{\hat{#1}}}}
\newcommand{\vecx}{\ensuremath{\vec{x}}}
\newcommand{\vecv}{\ensuremath{\vec{v}}}
\newcommand{\vech}{\ensuremath{\vec{H}}}
\newcommand{\vecj}{\ensuremath{\vec{J}}}
\newcommand{\vecp}{\ensuremath{\vec{p}}}
\newcommand{\vecn}{\ensuremath{\vec{n}}}
\newcommand{\veco}{\ensuremath{\vec{\Omega}}}
\newcommand{\veca}{\ensuremath{\boldsymbol\theta}}
\newcommand{\df}{\ensuremath{f}}
\newcommand{\tdf}{\ensuremath{\mathrm{tDF}}}
\newcommand{\paramsdiff}{\ensuremath{\vec{p}_\Delta}}
\newcommand{\paramsdf}{\ensuremath{\vec{p}_{\mathrm{DF}}}}
\newcommand{\paramspot}{\ensuremath{\vec{p}_\Phi}}
\newcommand{\sigv}{\ensuremath{\sigma_v}}

\newcommand{\DF}{\df}
\newcommand{\dff}{\ensuremath{\mathbf{f}}}
\newcommand{\dex}{\ensuremath{\,\mathrm{dex}}}
\newcommand{\Myr}{\ensuremath{\,\mathrm{Myr}}}
\newcommand{\Gyr}{\ensuremath{\,\mathrm{Gyr}}}
\newcommand{\kpc}{\ensuremath{\,\mathrm{kpc}}}
\newcommand{\pc}{\ensuremath{\,\mathrm{pc}}}
\newcommand{\kms}{\ensuremath{\,\mathrm{km\ s}^{-1}}}
\newcommand{\msun}{\ensuremath{\,\mathrm{M}_{\odot}}}
\newcommand{\inv}{\ensuremath{^{-1}}}

\newcommand{\jr}{\ensuremath{J_R}}
\newcommand{\jphi}{\ensuremath{J_\phi}}
\newcommand{\jz}{\ensuremath{J_Z}}
\newcommand{\lz}{\ensuremath{L_Z}}
\newcommand{\Or}{\ensuremath{\Omega_R}}
\newcommand{\Ophi}{\ensuremath{\Omega_\phi}}
\newcommand{\Oz}{\ensuremath{\Omega_Z}}
\newcommand{\ar}{\ensuremath{\theta_R}}
\newcommand{\aphi}{\ensuremath{\theta_\phi}}
\newcommand{\az}{\ensuremath{\theta_Z}}

%\submitted{}

\begin{document}

\title{Dynamical modeling of tidal streams}
\author{Jo~Bovy\altaffilmark{\ref{Hubble}}}
\affil{Institute for Advanced Study, Einstein Drive, Princeton, NJ 08540, USA}
\email{bovy@ias.edu}
\altaffiltext{\theaddress}{\label{Hubble}\stepcounter{address} Hubble fellow}

\begin{abstract} 
  I present a new framework for modeling the dynamics of cold tidal
  streams. The framework consists of simple models for the initial
  action--angle distribution of tidal debris as a function of time,
  which can be straightforwardly evolved forward in time. Taking
  advantage of the essentially one-dimensional nature of cold tidal
  streams, the transformation to position--velocity coordinates can be
  linearized and interpolated near a small number of points along the
  progenitor orbit, thus allowing for efficient computations of a
  stream's properties in observable quantities. I illustrate how to
  calculate the stream's track in different coordinate systems, how to
  quickly estimate the dispersion perpendicular to the track, and how
  to draw mock stream data. As a generative model, this framework
  allows one to compute the full probability distribution function and
  marginalize over or condition it on certain phase--space dimensions
  as well as convolve it with observational uncertainties. This will
  be instrumental in proper data analysis of stream data. In addition
  to providing a computationally-efficient practical tool for modeling
  the dynamics of tidal streams, the action--angle nature of the
  framework helps elucidate in exactly what manner streams do not
  follow single orbits, how the observed width of the stream relates
  to the velocity dispersion or mass of the progenitor, and how the
  progenitor of ``orphan'' streams could be located.

  The practical usefulness of the proposed framework crucially depends
  on the ability to calculate action--angle variables for any orbit in
  any gravitational potential. A novel method for calculating actions,
  frequencies, and angles in any static potential using a single orbit
  integration is described in an Appendix.
\end{abstract}

\keywords{
        dark matter
        ---
	Galaxy: halo
	---
        Galaxy: kinematics and dynamics
        ---
	Galaxy: structure
        ---
        galaxies: interactions
        ---
        stellar dynamics
}

\section{Introduction}

Tidal streams hold enormous promise as probes of both the large-scale
structure of the Milky Way (MW) halo's density distribution
\citep[\eg,][]{Johnston99a,Koposov10a} and its small-scale
fluctuations \citep{Carlberg09a,Carlberg12a}. However, two factors
have hampered the practical use of tidal streams in obtaining
constraints on the MW gravitational potential. First, streams do not
trace a single orbit, which has led to confusion over how to best fit
streams and over how problematic the single-orbit approximation really
is \citep{Eyre11a,Sanders13a}. Second, approaches that go beyond the
single-orbit assumption encounter practical difficulties of
computational cost and observational data quality making many of these
approaches impractical for real, noisy data
\citep{Sanders13b,PriceWhelan13a}. Observed gaps in tidal streams may
be due to interactions with dark--matter subhalos \citep{Yoon11a}, but
spurious underdensities could be created by the dynamics of stream
stars. This hinders our ability to use underdensities in star counts
along a tidal stream in order to constrain the number of encounters
with dark satellites and their masses \citep{Ngan14a}. Fast,
generative models of tidal streams would greatly benefit both of these
applications.

It has long been clear that the dynamics of tidal streams is most
simply described in terms of action--angle coordinates
\citep{Tremaine99a,Helmi99a}. Once a star has been tidally stripped
from the progenitor, the self-gravity of the stream can be neglected
and the orbital actions of a stream member are conserved while the
angles increase linearly with time. The action--angle structure of a
stream is therefore characterized by strong but straightforward
correlations between the actions and angles of stream members, the
exploitation of which is crucial for using streams to measure the host
gravitational potential \citep{Sanders13b}. The description of a
stream in action--angle coordinates also elucidates the connection
between the orbit and velocity dispersion (or mass) of the progenitor
and the action distribution of the tidal debris \citep{Eyre11a},
allowing simple physical models of the stream to be used and
constrained by observational data.

However, streams are observed in position--velocity coordinates and
for action--angle descriptions to be useful, we must be able to
calculate the transformation between these coordinate systems
efficiently. Until now this has required accurate phase--space data
\citep{Sanders13b} and specialized algorithms to calculate
action--angle coordinates that break down for the eccentric orbits on
which tidal-stream progenitors are typically found (that is, radial
and/or vertical actions of similar size as the angular momentum;
\citealt{Sanders12a}). However, even with Gaia, future stream data
will typically have large uncertainties compared to the intrinsic
dispersion of the stream and in particular the line-of-sight
velocities of the faintest stream members will not be observed for
many stars. Additionally, streams are superimposed on a non-negligible
background of field stars and background contamination cannot be
easily taken into account in any of the current stream-fitting
methods.

In this paper, I present a new method for modeling cold tidal streams
that fundamentally lives in frequency--angle space, because after
specifying the time at which a star was stripped, the stream
distribution function is essentially one-dimensional in this
space. While many of the results of this paper are more generally
valid, the fiducial stream model in this paper consists of a
three-dimensional (close to) Gaussian distribution of frequencies in
the stream, a three-dimensional Gaussian distribution of initial angle
offsets between stream members and the progenitor, and a uniform
distribution of stripping times; the frequency and angle distributions
are independent of stripping time. While this model oversimplifies the
real stripping process, it is argued that at least for constraining
the gravitational potential with stream data, this simple model is
likely unbiased. For detailed mock stream data, more elaborate models
of the stripping process as a function of time might be necessary and
I discuss how these could easily be incorporated into the proposed
framework.

To evaluate the model in position--velocity coordinates, the
transformation to and from action--angle coordinates is computed using
a novel method for calculating action--angle coordinates presented in
Appendix~\ref{sec:aa}. This transformation is linearized at a small
number of points along the progenitor orbit that span the width of the
stream. For cold streams, the separation between the stream track and
the progenitor orbit as well as the internal differences within the
stream are small, such that this linear approximation is accurate
enough. The linear approximation allows for fast evaluation of the
average stream location (the stream ``track'') in position--velocity
coordinates and efficient mock data generation. For a uniform
distribution of stripping times, an observed stream member's
phase--space probability distribution function (PDF) can be
analytically marginalized over stripping time, such that using the
linear approximation any evaluation---including marginalization and
uncertainty convolution---of the PDF is extremely rapid. This clears
the way for proper probabilistic inference of the Milky Way halo's
gravitational potential and progenitor properties using stream data
for individual stars.

This paper is outlined as follows. In \sectionname~\ref{sec:dynamics}
I briefly summarize the dynamics of tidal streams in terms of
action--angle coordinates. A generative model of a tidal stream in
frequency--angle coordinates is given in
\sectionname~\ref{sec:generative}. \sectionname~\ref{sec:track}
describes how to compute the stream's properties as a function of
angle along the stream, both in action--angle coordinates and in
poition--velocity coordinates. \sectionname~\ref{sec:mock} discusses
how to generate mock stream data using the generative model of a tidal
stream and in \sectionname~\ref{sec:pdf} I show how to calculate the
stream PDF for individual stream members as well as its
marginalization and convolution over missing and noisy directions. A
discussion and outlook is presented in
\sectionname~\ref{sec:discussion} and I conclude in
\sectionname~\ref{sec:conclusion}. In \appendixname~\ref{sec:aa} a
novel method for calculating actions, frequencies, and angles for any
orbit in any static gravitational potential is presented.

%BOVY: reference point

\section{The dynamics of tidal streams}\label{sec:dynamics}

The dynamics of tidal streams both in real (position--velocity) and
action--angle space has been described by various previous authors
\citep[\eg,][]{Helmi99a,Tremaine99a,Johnston99a,Sanders13b}. I provide
here a brief description of the dynamics of tidal streams that is
relevant for what follows.

In real space, a tidal stream forms as stars are stripped from a
progenitor some time $\Delta t$ in the past and then evolve (largely)
independently in a galaxy's host potential, moving away both in
position and velocity from the progenitor object as time goes
on. Thus, a star was offset by a small amount $(\Delta \vecx,\Delta
\vecv)$ from the position of the progenitor $(\vecx^p,\vecv^p)(-\Delta
t)$ at that time. Then the star orbited under the influence of the
same gravitational potential $\Phi$ as the progenitor until the
present day $(t=0)$, when it was observed: $(\vecx,\vecv)(t=0)$ =
$\vech_{\Delta t}(\vecx^p(-\Delta t)+\Delta \vecx,\vecv^p(-\Delta
t)+\Delta \vecv)$, where $\vech$ denotes the Hamiltonian flow from
$t=-\Delta t$ to $t=0$ (see \citealt{binneytremaine}); this is simply
the orbit of the star between these times. In phase-space this orbit
can be calculated by solving Hamilton's equations: $\dot{\vecx} =
\vecv; \dot{\vecv} = - \dd \Phi / \dd \vecx$.

In action--angle space the dynamics of stream formation is much
simpler. A star received an offset ($\Delta \vecj,\Delta \veca$) at
time $-\Delta t$ from the action--angle coordinates of the progenitor
$(\vecj^p,\veca^p)$, similar to the small offset ($\Delta \vecx,\Delta
\vecv)$ in the real--space description. The offset in the actions
corresponds to an offset in the frequencies $\Delta \veco$;
ultimately, this frequency offset is responsible for the spreading of
the stream over long stretches on the sky. Because both the
progenitor's angles and the star's angles increase linearly with time,
albeit at different frequencies, the difference in the angles also
increases linearly in time. Therefore at time $t=0$ the difference in
angles is
\begin{equation}
\Delta \veca(t=0) = \Delta \veca(-\Delta t) + \Delta \veco \Delta t\,,
\end{equation}
while the difference in actions is constant
\begin{equation}
\Delta \vecj(t=0) = \Delta \vecj(-\Delta t)\,.
\end{equation}
As the initial offset $\Delta \veca(-\Delta t)$ in angles is small
compared to the growth, we have that $\Delta\veco\Delta t$, $\Delta
\veca(t=0) \approx \Delta \veco \Delta t$.

Thus, in action--angle coordinates, the dynamics of tidal-stream
formation is simplified to linear growth of small initial angle
differences due to small action/frequency differences. The angle
direction $\unitvec{\veco} \equiv \Delta \veco / | \Delta \veco|$
along which a star moves away from the progenitor is related to the
initial action offset by $\Delta \veco \approx (\partial^2 H /
\partial \vecj \partial \vecj) \Delta \vecj$, in the limit of small
action offsets, because $\veco = \partial H / \partial \vecj$. For an
approximately one-dimensional stream to form, the Hessian matrix
$(\partial^2 H / \partial \vecj \partial \vecj)\large|_{\vecj^p}$ has
to be dominated by a single large eigenvalue; in this case the angle
difference of stars stripped from the stream at any time will fall
along approximately the same direction $\unitvec\veco$, and the stream
will essentially be one dimensional. In detail, because the
distribution of initial offsets $\Delta \vecj$ is not isotropic, the
direction $\unitvec \veco$ is not just determined by the potential
(through $H$) and the progenitor orbit, but also by the distribution
of $\Delta \vecj$, which is set through a combination of the internal
dynamics of the progenitor and the stripping process (see discussion
in \citealt{Sanders13b}). For dynamical modeling we can be agnostic
about the relative importance of all of these contributors to
$\unitvec \veco$ by flexible modeling of the distribution of $\Delta
\vecj$ or $\Delta \veco$.

\section{Generative models of tidal streams}\label{sec:generative}

\subsection{A simulation}\label{sec:sim}

I illustrate the theoretical considerations set forth in this paper by
using a simulated tidal stream. This stream is set up similarly to
that used in \citet{Sanders13b}. The stream is constructed by running
an $N$-body simulation of a King cluster on an orbit similar to that
of the GD-1 stream \citep{Koposov10a}. Specifically, a King cluster
with a mass of $2\times10^4\msun$, a tidal radius of $0.07\kpc$, and a
ratio of the central potential to the velocity dispersion squared of
$2.0$ is sampled using $10^4$ particles. The cluster is evolved in an
external logarithmic host potential with $V_c = 220\kms$ and a
flattening $q=0.9$ ($\Phi = V_c^2 \ln [R^2 + z^2/q^2]$) using the
gyrfalcon code \citep{Dehnen00a,Dehnen02a} with a softening of
$1.5\pc$ in the NEMO toolkit \citep{Teuben95a}. The cluster is evolved
for $5.011\Gyr$ ($=5.125\kpc/[\kms]$) from the initial condition
$(X,Y,Z) = (-11.63337239,-20.76235661,-10.631736273934635)\kpc$ and
$(V_X,V_Y,V_Z) = (-128.8281653,42.88727925,79.172383882274971)\kms$,
chosen such that the cluster ends up at roughly the observed location
of the GD-1 stream at the end of the simulation
\citep{Koposov10a}. The progenitor's orbit has an eccentricity of
$0.31$, a pericenter of $13.7\kpc$, a current $r=14.4\kpc$, such that
the cluster is close to pericenter at the end of the simulation. The
actions of the progenitor are $(J_R,L_Z,J_Z) =
(288.5,-3173.7,897.6)\kpc\kms$ and its orbital frequencies are
$(\Omega_R,\Omega_\phi,\Omega_Z) = (15.7,-10.8,11.9)\Gyr\inv$.

The final position of the stream in $(X,Z)$ is shown in
\figurename~\ref{fig:gd1_xz}. \figurename~\ref{fig:gd1_jt} shows the
final position of the stream in action--frequency--angle coordinates,
calculated using the algorithm described in
\appendixname~\ref{sec:aa}. As described in
\sectionname~\ref{sec:dynamics}, a one-dimensional tidal stream forms
because even though the action distribution of the tidal debris is not
isotropic or one-dimensional, the Hessian $(\partial^2 H / \partial
\vecj \partial \vecj)\large|_{\vecj^p}$ is so strongly dominated by a
single large eigenvalue (the ratio of the largest to the second
largest eigenvalue is about $30$), that the frequency distribution of
the debris is essentially one dimensional. The angle distribution of
the debris at the end of the simulation is therefore essentially one
dimensional as well.

\subsection{A model in frequency--action space}\label{sec:modeloa}

The action and frequency distributions shown in
\figurename~\ref{fig:gd1_jt} combined with the knowledge that for
\emph{any} tidal stream the Hessian $(\partial^2 H / \partial \vecj
\partial \vecj)\large|_{\vecj^p}$ must be strongly dominated by a
single large eigenvalue leads us to propose that frequency--angle
space is a better coordinate system than action--angle space for
modeling tidal streams. That is, the distribution of actions in a
tidal stream has a complicated structure that is hard to capture using
a simple, analytic form (see also the discussion in
\citealt{Eyre11a}). Especially the ``bowtie'' structure in $(J_R,L_Z)$
and the overlapping leading- and trailing-arm distributions in
($J_R,J_Z$) are hard to model with a simple distribution function. The
distribution of frequencies, however, is close-to one dimensional, its
direction of largest variance can be well-modeled as a Gaussian
(although below we will model it slightly differently), and its
parameters can all be easily estimated from the velocity dispersion of
the progenitor, the orbit of the progenitor, and the gravitational
potential, as discussed below. Therefore, we will model tidal streams
in frequency--angle space. Because the leading and trailing arm are
well-separated in frequency space, we model each arm individually.

A generative model of a tidal stream in frequency--angle space
requires three ingredients: (a) a model for the distribution of times
at which stars are stripped from the progenitor, (b) a prescription
for the distribution of frequency offsets from the progenitor at every
given stripping time, and (c) a description of the angle offsets for
any given frequency offset and stripping time. Having specified these
ingredients, we can then generate tidal debris at any given stripping
time and evolve it forward using the simple linear dynamics in
frequency--angle space discussed in
\sectionname~\ref{sec:dynamics}. The initial angle offsets are small
enough that they are likely unobservable even with futuristic data
(especially in the direction of the frequency offset, where the
initial angle offset is quickly overwhelmed by the subsequent
dynamical evolution; for the simulated stream used here this happens
after approximately $20\Myr$). For that reason, we will assume that
the initial angle offsets are independent of both stripping time and
frequency offset. In what follows, we will model them using a simple
isotropic Gaussian distribution.

Ingredients (a) and (b) require more careful modeling. The right panel
of \figurename~\ref{fig:gd1_times} shows the distribution of times at
which stream particles are stripped from the progenitor. These
stripping times are calculated from the final snapshot by linear
regression of a particle's angle offset $\Delta \veca$ from the
progenitor, versus the frequency offset $\Delta \veco$ as 
\begin{equation}
  \Delta t = \frac{\Delta \veco \cdot \Delta \veca}{|\Delta \veco|^2}\,.
\end{equation}
The dotted line shows a finely-binned histogram of these stripping
times, which indicates that stripping happens in bursts with a radial
period of about $400\Myr$, the radial period of the progenitor's
orbit. Thus, as expected, stripping happens primarily at pericenter
passages, with a smaller number of particles lost between pericenter
passages. Realistic models of a stream, especially those that require
good models of the surface-density structure of the stream as a
function of stream angle, need to model this non-uniform stripping
process.

The middle panel of \figurename~\ref{fig:gd1_times} shows the
frequency offsets between stream particles and the progenitor as a
function of strippng time, corresponding to ingredient (b) above. The
same burstiness that is apparent in the distribution of stripping
times shows up here. Particles stripped at pericenter typically have
larger frequency offsets than those lost at larger radii. This can be
described as saying that particles are \emph{removed} at pericenter,
but only \emph{peeled off} between pericenter passages. Again,
realistic models of tidal streams need to take this detailed structure
into account.

The right panel of \figurename~\ref{fig:gd1_times} shows the angle
offset (at the end of the simulation) versus the stripping time. There
is a spread in the angle offset reach for any given stripping time,
but it is nevertheless typically the case that particles that have
been stripped earlier are found at larger angle offsets
today. Self-sorting---the tendency of particles stripped later with
larger frequency offsets to overtake particles stripped earlier at
smaller frequency differences and thus to erase correlations between
particle position along the stream and stripping time---happens to a
small extent but the narrow distribution of frequency offsets limits
the ability of particles stripped later to reach large distances from
the progenitor.

\subsection{The fiducial stream model}

The discussion of the simulated stream's properties as a function of
stripping time in the previous section indicates that we can build
simple models of the structure of tidal streams when considering them
in frequency--angle space. In this section, we propose a very simple
fiducial model that we will use in the remainder of this paper. This
model combines simple analytic forms for ease of computation that are
nevertheless realistic enough to provide adequate models for tidal
streams in many applications.

First, we approximate the bursty, non-uniform distribution of
stripping times in \figurename~\ref{fig:gd1_times} with a uniform
distribution
\begin{align}\label{eq:pt}
  \begin{split}
    p(\Delta t) & \propto 1\,,\qquad 0 < \Delta t < t_d\\
    & = 0\,,\qquad \mathrm{otherwise}\,.
  \end{split}
\end{align}
The thick, solid histogram in the right panel of
\figurename~\ref{fig:gd1_times} shows a coarser-binned histogram of
stripping times, indicating that the uniform distribution is a good
approximation over intervals larger than a radial period ($\approx
400\Myr$). The \emph{disruption time} $t_d$ is a free parameter of the
model that needs to be determined for each stream. For the model of
the simulated stream used here we set $t_d = 4.5\Gyr$.

Similarly, we model the distribution of frequency offsets between
stream particles and the progenitor to be independent of stripping
time, even though the middle panel of \figurename~\ref{fig:gd1_times}
clearly shows that particles are stripped at larger frequency
differences at pericentric passages than between them. The fact that
there is no long-term trend in the typical frequency offset means that
this is a good model on time-scales larger than the radial
period. While this means that the small-scale structure of the stream
will not be perfectly represented by the model, the large-scale
structure of the stream will be fine. For using tidal streams to
constrain the gravitational potential, large arcs are most important,
such that this simple model will be adequate for such inferences.

The distribution of frequencies is three dimensional. Because we model
the leading and the trailing arm separately, a single-peaked
distribution suffices to describe the frequency distribution. A
general description in terms of a Gaussian distribution would have
nine free parameters. However, we can model it with fewer free
parameters as follows. First, following \citet{Eyre11a}, we
approximate the \emph{action} distribution as a Gaussian with standard
deviations
\begin{equation}
  \sigma_{J_R} = \frac{1}{\pi}\,\sigv\,\left(r_{\mathrm{apo}}-r_{\mathrm{peri}}\right)\,,\quad 
  \sigma_{L_Z} = \sigv\,r_{\mathrm{peri}}\,,\quad 
  \sigma_{J_Z} = \frac{2}{\pi}\,\sigv\,Z_{\mathrm{max}}\,,
\end{equation}
where $\sigv$ is the velocity dispersion of the progenitor,
$r_{\mathrm{apo}}$ and $r_{\mathrm{peri}}$ are the apo- and pericenter
radii of the progenitor orbit, respectively, and $Z_{\mathrm{max}}$ is
the maximum $Z$ reached on this orbit. We assume that the all
correlations between the actions are zero (see
\figurename~\ref{fig:gd1_jt}). Then we propagate this Gaussian
variance to frequency space using the Hessian $(\partial^2 H /
\partial \vecj \partial \vecj)\large|_{\vecj^p}$\footnote{The Hessian
  $(\partial^2 H / \partial \vecj \partial \vecj)\large|_{\vecj^p}$
  can be computed using the method in \appendixname~\ref{sec:aa} by
  computing $\partial (\vecj,\veca) / \partial (\vecx,\vecv)$ and
  $\partial (\veco,\veca) / \partial (\vecx,\vecv)$ and forming
  $\partial \veco / \partial \vecj = \left(\partial (\vecj,\veca) /
  \partial (\vecx,\vecv)\right)^{-1}\,\left(\partial (\veco,\veca) /
  \partial (\vecx,\vecv)\right)$.}, which gives the variance matrix
$V_\veco$. The principal eigenvector of this matrix is the model
direction along which the stream spreads. While the model I propose
here is a simple one, it works well without having been tweaked for
the simulation under scrutiny. The angle between the principal
eigenvector of $V_\veco$ and the progenitor's frequency vector is
$0.50^\circ$ while the value measured from the simulation is
$0.18^\circ$. For a model with an isotropic action distribution, the
misalignment would be $1.28^circ$.

We then further constrain the mean frequency offset to lie along the
principal eigenvector of $V_\veco$. This mean frequency offset can be
described in terms of a parameter $\mu_\Omega = \Delta \Omega^m /
\sigma_{\Omega,1}$, where $\sigma_{\Omega,1}$ is the square root of
the largest eigenvalue of $V_\veco$ and $\Delta \Omega^m$ is the mean
frequency offset. The sign of this parameter sets whether we are
modeling a leading or a trailing arm. \figurename~\ref{fig:gd1_dohist}
shows the distribution frequency offsets in the direction of the
median frequency offset in the stream. The dashed line is a Gaussian
fit to this distribution, while the solid line is a fit of the form
\begin{equation}\label{eq:do}
  p(|\Delta \veco|_\parallel) = |\Delta \veco|_\parallel\,\mathcal{N}\left(|\Delta \veco|_\parallel|\Delta \Omega^m,\sigma_{\Omega,1}^2\right)\,,
\end{equation}
where $\mathcal{N}(\cdot|\cdot,\cdot)$ represents a Gaussian
distribution. The fit is equally good (because the dispersion is much
smaller than the mean) and we choose the second form because it
simplifies some of the stream-track calculations below. The best-fit
$(\Delta \Omega^m,\sigma_{\Omega,1})$ is $(0.19,0.033)\Gyr\inv$ in
both cases. Based on this information, a model with $\sigv =
0.365\kms$ and $\mu_\Omega = 6$ provides a good fit to the simulation
data. The distribution of frequency offsets in the two-dimensional
space perpendicular to the principal eigenvector of $V_\veco$ is
modeled as a zero-mean Gaussian with variance matrix given by the
projection of $V_\veco$ onto this space.

Because we expect the typical $\Delta\veco$ to scale with the
velocity dispersion of the progenitor (see also \citealt{Sanders13a}),
I conjecture that a constant $\mu_\Omega$ can be used for modeling
cold tidal streams of any (small-ish) velocity dispersion. Further
modeling of simulated streams, however, is necessary to check this and
to find the optimizal value of $\mu_\Omega$. Fixing $\mu_\Omega$
leaves a single free parameter \sigv\ for modeling the frequency
offset distribution of a tidal stream. In modeling to observed data,
this parameter needs to be fit and we expect it to be proportional to
the velocity dispersion of the progenitor, although this
proportionality needs to be checked more carefully (but see Figure 3
in \citealt{Sanders13a}, which shows that the size of the frequency
distribution scales as mass$^{1/3}$, or approximately as the velocity
dispersion). Similarly, the initial angle distribution, modeled here
as an isotropic Gaussian, has a charactaristic spread $\sigma_{\theta}$
which we set here to $0.00365$ ($=\sigv/[100\kms]$), and we likewise
expect this spread to scale with the velocity dispersion of the
progenitor.

The fiducial model for a tidal stream in frequency--angle coordinates
he proposed here is therefore described by essentially two free
parameters in addition to the progenitor's phase-space position: the
progenitor's velocity dispersion \sigv\ and the disruption time
$t_d$. As shown below, these two parameters are the most important in
determining the track of the stream. More sophisticated models can
leave $\mu_\Omega$ and $\sigma_{\theta}$ as free paramters as well. In
what follows, we will fix these to the values given in the previous
paragraph. It must be stressed, however, that all of these parameters
here have been fixed ``by-eye'', and better fits might be possible
using more quantitative fit procedures. But such procedures are not
pursued here further.

\section{The track of a tidal stream}\label{sec:track}

\section{Mock stream data}\label{sec:mock}

\section{The full stream PDF}\label{sec:pdf}

\section{Discussion}\label{sec:discussion}

\section{Conclusion}\label{sec:conclusion}

\subsection{Action--angle stream modeling}

Similar as for the real--space approach, we can write down a
generative model for the positions and velocities ($\vecx_i,\vecv_i$)
of stars $i$ in a tidal stream in the action--angle formalism by
specifying: (a) the gravitational potential $\Phi$, (b) the current
action-angle coordinates of the progenitor $(\vecj_p,\veca_p)(t=0)$,
(c) the time $\Delta t_i$ at which each star was removed from the
progenitor, and (d) the distribution $f(\Delta \vecj,\Delta
\veca|\paramsdiff)$, characterized by some parameters $\paramsdiff$,
of initial phase-space offsets from which the actual offsets $(\Delta
\vecj_i,\Delta \veca_i)$ are drawn. In practice, $f$ can again be
taken to be a Gaussian, although more general models could take the
expected `bowtie' structure of the action distribution into account
(\citealt{Eyre11a}; see also \figurename~\ref{fig:gd1_jt}).

The probability for observing a star today at $(\vecx,\vecv)$ given
items (a) through (d) is then
\begin{align}\label{eq:aadft}
  p(\vecx_i,\vecv_i | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) 
  & = p(\vecj_i,\veca_i | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,,\nonumber\\
  & = p(\vecj_i-\vecj_p,\veca_i-\veca_p-\Delta \veco_i\Delta t_i | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,,\\
  & = f(\Delta \vecj_i,\Delta \veca_i(-\Delta t_i) | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,,\nonumber
\end{align}
where the first equality holds because the Jacobian of the
$(\vecx,\vecv)\rightarrow(\vecj,\veca)$ transformation is unity and
the second equality is a simple linear transformation. The final
probability can again be evaluated using the parameters
\paramsdiff\ following (d).

As in the real--space formalism, we should integrate
\eqnname~(\ref{eq:aadft}) over the time $\Delta t_i$ at which the star
is removed from the progenitor (cf.~\eqnname~[\ref{eq:realdf}]):
\begin{equation}\label{eq:aadf}
  p(\vecx_i,\vecv_i | \Phi,\vecj_p,\veca_p,\paramsdiff) = \int \dd
  \Delta t_i\, f(\Delta \vecj_i,\Delta \veca_i(-\Delta t_i) |
  \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,.
\end{equation}
Without further assumptions the action-angle formalism proceeds in the
same way as the real--space formalism, evaluating
\eqnname~(\ref{eq:aadft}) for a grid $j$ of times $\Delta t_i^j$ and
summing $f(\Delta \vecj_i,\Delta \veca_i(-\Delta t^j_i) |
\Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff)$ over $j$.

Even without further assumptions about the shape of $f(\Delta
\vecj,\Delta \veca|\paramsdiff)$, the action-angle formalism already
allows for a simplification over the real--space method. Because the
progenitor orbit is specified in action-angle coordinates, it can be
changed \emph{without} expensive orbit integrations of all possible
source orbits. As the progenitor's current phase-space position is a
six-dimensional parameter, this is a major simplification of the
dynamical model when the source orbit is unknown. Note that this
simplification does require us to be able to compute $\veco(\vecj)$,
to calculate the new source frequencies $\veco_p$ when changing the
source actions $\vecj_p$. This mapping can be constructed using a
small number of (action,frequency) calculations covering the range of
possible source actions and is relatively easy to construct due to the
approximately one-dimensional structure of a stream (see discussion at
the end of \sectionname~\ref{sec:dynamics}).

We can simplify \eqnname~(\ref{eq:aadf}) by assuming (a) that
$f(\Delta \vecj,\Delta \veca|\paramsdiff)$ is separable into
$f_{\veca}(\Delta \veca|\paramsdiff)\,f_{\vecj}(\Delta
\vecj|\paramsdiff)$ and (b) that the distribution $f_{\veca}(\Delta
\veca|\paramsdiff)$ is Gaussian. In what follows, we will assume that
this Gaussian is isotropic with dispersion $\sigma_\theta$, but all of
the results are the same after appropriate diagonalization and
rescaling of the angles for more general covariance matrices. In this
case, \eqnname~(\ref{eq:aadf}) can be written as
\begin{equation}
\begin{split}
  p(\vecx_i,\vecv_i & | \Phi,\vecj_p,\veca_p,\paramsdiff) 
   = \\
   & f_{\vecj}(\Delta \vecj_i|\paramsdiff) \int \dd \Delta t_i\,
  \frac{1}{(2\,\pi)^{3/2}\,\sigma_\theta^3}\,\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\left(\Delta \veca_i-\Delta \veco_i\Delta t_i\right)^2\right]\right)\,,
\end{split}
\end{equation}
where the sum in the argument of the exponential is over the three
components of $\Delta \veca_i$ and $\Delta \veco_i$. This integral can
be done analytically, resulting in
\begin{equation}\label{eq:aadfsimple}
\begin{split}
  p(\vecx_i,\vecv_i & | \Phi,\vecj_p,\veca_p,\paramsdiff) 
   = \\
   & f_{\vecj}(\Delta \vecj_i|\paramsdiff) \,  \frac{1}{2\,\pi\,\sigma_\theta^2\,\sqrt{\sum_{R,\phi,Z}\Delta \veco_i^2}}\,\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\Delta \veca^2_i-\frac{\left(\sum_{R,\phi,Z} \Delta \veco_i\Delta \veca_i\right)^2}{\sum_{R,\phi,Z}\Delta \veco_i^2}\right]\right)\,.
\end{split}
\end{equation}
In this expression, $\Delta X^2 \equiv (X_1-X_2)^2$. The expression in
square brackets is positive by using the Cauchy-Schwarz inequality.

The quantity in the square brackets in \eqnname~(\ref{eq:aadfsimple})
expresses the fact that the present-day difference in angles between a
star in a tidal stream and the source must be linearly related by a
scalar (the time since the removal of the star) to the difference in
frequencies, within a tolerance set by the width of the distribution
of initial angle differences ($\sigma_\theta$). The ratio
$\sum_{R,\phi,Z} \Delta \veco_i\Delta \veca_i/\sum_{R,\phi,Z}\Delta
\veco_i^2$ is the best-guess for $\Delta t_i$ when performing a linear
regression between $\Delta \veca_i$ and $\Delta \veco_i$. The
expression in square brackets is minimized when the correlation
between $\Delta \veca_i$ and $\Delta \veco_i$ is the largest, \ie,
when a single time difference $\Delta t_i$ explains all three of the
angle differences.

A detail that is ignored in \equationname~(\ref{eq:aadfsimple}) is
that the uniform prior on $\Delta t$ should only have support for
$\Delta t > 0$, \ie, for times in the past, which introduces an
additional factor of $[1+\mathrm{erf}(a_0)]$, with
\[ a_0 = \frac{\sqrt{\sum_{R,\phi,Z}\Delta
  \veco_i^2}}{\sqrt{2}\,\sigma_\theta}\,\frac{\sum_{R,\phi,Z} \Delta
  \veco_i\Delta \veca_i}{\sum_{R,\phi,Z}\Delta \veco_i^2}\,.\] Because
the second factor on the right-hand side is the best guess for $\Delta
t_i$, if this is large and negative (\ie, the star was stripped
\emph{in the future}), the probability of the star being stripped from
the progenitor is strongly suppressed. Only when it appears that the
star was only recently stripped with $\Delta t_i$ negative, but
$\approx \sigma_\theta/\sqrt{\sum_{R,\phi,Z}\Delta \veco_i^2}$, is the
suppression for negative stripping times less.

The discussion so far has been entirely general and applies to any
sort of stellar debris that originates from a single orbit, without
necessarily generating a one-dimensional structure. In most
observations of tidal streams, however, they appear as one-dimensional
objects rather than the six-dimensional objects that we have
considered so far. The reason for this is both physical and
observational: (a) narrow tidal streams form because a star separates
from the progenitor over time mostly along a single direction in
angle-space: the Hessian matrix of the Hamiltonian, $\partial^2 H /
\partial \vecj \partial \vecj$, has a single large eigenvalue, which
determines the direction along which the stream is disrupted in
angle-space (because $\partial \veco / \partial \vecj$ is equal to the
Hessian); (b) because many tidal streams are dynamically cold with
velocity dispersions $\lesssim10\kms$, observations are unable to
resolve the angle differences along the two eigenvectors of the
Hessian corresponding to the two smallest eigenvalues or the spread in
actions among stream members. If the latter can be resolved, the
stream is four-dimensional. In what follows we will be mainly
concerned with the stream only spreading along a single direction in
angle space.

If the stream is one-dimensional in angle space then the time since
the removal from the progenitor becomes unobservable if the progenitor
orbit is unknown: above, determining the time depended on explaining
the three-dimensional angle difference $\Delta \veca$ with the
three-dimensional frequency difference $\Delta \veco$ using a single
time difference; if two of the angle and frequency differences are
unable to be resolved, then the frequency difference and the time
difference become degenerate (aka, you can draw a straight line
connecting any two points). To make use of this fact to simplify the
dynamical model, we write the frequency difference as
\begin{equation}
  \Delta \veco = \Delta \Omega_1\,\unitvec{e}_1\,,
\end{equation}
where $\unitvec{e}_1$ is a unit vector. Then we write the angle
difference $\Delta \veca$ 
\begin{equation}
  \Delta \veca = \Delta \theta_1\,\unitvec{e}_1 + \Delta \veca_\perp\,\unitvec{e}_1^\perp\,,
\end{equation}
where$\unitvec{e}_1^\perp$ is the orthogonal complement of
$\unitvec{e}_1$ (the two directions perpendicular to $\unitvec{e}_1$)
and $\Delta \veca_\perp$ are the components of $\Delta \veca$ in that
direction. We can then write \eqnname~(\ref{eq:aadfsimple}) as
\begin{equation}\label{eq:aadfsimpler}
\begin{split}
  p(\vecx_i,\vecv_i & | \Phi,\vecj_p,\veca_p,\paramsdiff) 
   = \\
   & f_{\vecj}(\Delta \vecj_i|\paramsdiff) \,  \frac{1}{2\,\pi\,\sigma_\theta^2\,\sqrt{\sum_{R,\phi,Z}\Delta \veco_i^2}}\,\exp\left(-\frac{|\Delta\veca_{i,\perp}|^2}{2\,\sigma_\theta^2}\right)\,.
\end{split}
\end{equation}
\Eqnname~(\ref{eq:aadfsimpler}) is equivalent to
\eqnname~(\ref{eq:aadfsimple}), but it expresses the dynamical model
in a slightly different way. Whereas the angle part of
\eqnname~(\ref{eq:aadfsimple}) allows the correct potential and
progenitor to be inferred by asking whether a single time difference
explains the three angle differences given the three frequencies,
\eqnname~(\ref{eq:aadfsimpler}) `rewinds' the star's orbit along the
direction of $\Delta \veco$ and asks whether the angle difference
perpendicular to $\Delta \veco$ is consistent with the initial angle
spread (because it has not evolved since the removal of the star). In
re-writing this, we have not yet used the approximation that the
stream only forms along a single direction in $\Delta \veco$.

\section{Two equivalent generative models of tidal streams}\label{sec:method}

In this Section, we describe two mathematically equivalent generative
models for a tidal stream. The first is based on orbit integrations in
real (position-velocity) space: a tidal stream is generated from small
(phase-space) offsets between a star and a source\footnote{The word
  `progenitor' is typically used to indicate the actual parent body of
  a star in a tidal stream. For the dynamical model only a reference
  orbit from which the tidal stream can be thought of to originate is
  necessary; this reference orbit does \emph{not} have to coincide
  with the actual progenitor (this point is most clear in the
  description of the dynamics in action-angle coordinates). To avoid
  confusion between the hypothetical progenitor and the actual parent
  body, we refer to the reference orbit as the `source' rather than
  the `progenitor'.} that are evolved in time using orbit
integrations. The second method relies on the equivalent description
of this process in action-angle coordinates. While these two methods
are equivalent, the latter provides a conceptually simpler description
of the dynamical model and, when fitting a tidal stream, it allows for
fast marginalization over nuisance parameters. In particular, the
marginalization over the time at which a star was removed from the
source is analytic and the integration over possible source orbits can
be performed much faster (by avoiding orbit integrations of all
possible source orbits).

\subsection{Real--space model}\label{sec:realmethod}

The description of the stripping of a star from a satellite in real
(\vecx,\vecv) space assumes that at a time $\Delta t$ ago the star was
offset by a small amount ($\Delta \vecx,\Delta \vecv$) from the
position of the source $(\vecx^s,\vecv^s)(-\Delta t)$ at that
time. Then the star orbited under the influence of the same
gravitational potential $\Phi$ as the source until the present day
$(t=0)$, when it was observed: $(\vecx,\vecv)(t=0)$ = $\vech_{\Delta
  t}(\vecx^s(-\Delta t)+\Delta \vecx,\vecv^s(-\Delta t)+\Delta
\vecv)$, where $\vech$ denotes the Hamiltonian flow from $t=-\Delta t$
to $t=0$ (see \citealt{binneytremaine}); this is simply the orbit of
the star between these times. In phase-space this orbit can be
calculated by solving Hamilton's equations: $\dot{\vecx} = \vecv;
\dot{\vecv} = - \dd \Phi / \dd \vecx$.

A generative model for the positions and velocities
($\vecx_i,\vecv_i$) of stars $i$ in a tidal stream is then fully
specified by: (a) the gravitational potential $\Phi$, (b) the current
phase-space position of the source $(\vecx_p,\vecv_p)(t=0)$, which can
be integrated backwards in time by reversing the Hamiltonian flow
defined by $\Phi$, (c) the time $\Delta t_i$ at which each star was
removed from the source, and (d) the distribution $f(\Delta
\vecx,\Delta \vecv|\paramsdiff)$, characterized by some parameters
$\paramsdiff$, of initial phase-space offsets from which the actual
offsets $(\Delta \vecx_i,\Delta \vecv_i)$ are drawn. In practice, $f$
can be taken to be a Gaussian, characterized by a width in position
$\sigma_x$ and in velocity $\sigma_v$, corresponding to the tidal
radius and escape velocity of the progenitor, respectively. However,
more complicated models for the initial phase-space distribution of
the debris can be used here as well (see discussion below). In the
case of a Gaussian $f$, the source can be thought of as the mean of
this Gaussian, with the additional requirement that this mean is an
orbit in the potential $\Phi$.

The probability for observing a star today at $(\vecx,\vecv)$ given
items (a) through (d) is then
\begin{align}\label{eq:realdft}
  p(\vecx_i,\vecv_i | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff) 
  & = p(\vecx_i(-\Delta t),\vecv_i(-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff) \,,\nonumber\\
  & = p([\vecx_i-\vecx_p](-\Delta t),[\vecv_i-\vecv_p](-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff) \,\nonumber\\
  & = p(\Delta\vecx_i(-\Delta t),\Delta\vecv_i(-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff)\,,\\
  & = f(\Delta\vecx_i(-\Delta t),\Delta\vecv_i(-\Delta t) | \Phi,\vecx_p,\vecv_p,\Delta t_i,\paramsdiff)\, ,\nonumber
\end{align}
where the first equality holds because the Hamiltonian flow is unique
and conserves phase-space volume by Liousville's theorem and the
second one because the Jacobian of the transformation
$(\vecx_i,\vecv_i)(-\Delta t) \rightarrow (\vecx_i,\vecv_i)(-\Delta t)
- (\vecx_p,\vecv_p)(-\Delta t)$ is one. The final probability can be
evaluated using the parameters $\paramsdiff$ following (d).

The time $\Delta t_i$ at which a star was removed is unobservable and
the probability in \eqnname~(\ref{eq:realdft}) should therefore be
integrated over this time using a prior $p(\Delta t_i|\ldots)$, where
$\ldots$ indicates that this prior could depend on other factors such
as the potential and the progenitor orbit\footnote{For example, one
  could insist that stars are removed around pericenter passages. Note
  that this assumption would require the source orbit to be the actual
  progenitor orbit.}. For simplicity, we will assume a uniform prior
for the times $\Delta t$. The probability for a single star then becomes
\begin{align}\label{eq:realdf}
  p(\vecx_i,\vecv_i | \Phi,\vecx_p,\vecv_p,\paramsdiff) & = \int \dd
  \Delta t_i \,p(\vecx_i,\vecv_i | \Phi,\vecx_p,\vecv_p,\Delta t_i,
  \paramsdiff)\,.
\end{align}
If the progenitor's current phase-space position is unknown, then this
probability should be further marginalized over $(\vecx_p,\vecv_p)$;
for multiple stars, this marginalization is done after multiplying the
individual-star likelihoods in \eqnname~(\ref{eq:realdf}).

In practice, the likelihood in \eqnname~(\ref{eq:realdf}) can be
calculated as follows: for a given potential $\Phi$ and source
$(\vecx_p,\vecv_p)$, we (a) integrate the orbits of all stars
$(\vecx_i,\vecv_i)$ and (b) that of the source backward in time for
many Gyr and record the phase-space positions over time on a grid $j$
in $\Delta^j_i$; (c) we evaluate $f(\Delta \vecx_i,\Delta
\vecv_i)(\Delta t^j_i)$ for all stars $i$ and times $j$; (d) $f(\Delta
\vecx_i,\Delta \vecv_i)(\Delta t^j_i)$ is then summed over $j$ and the
logarithm of that is summed over $i$. The most computationally
expensive part of this calculation is the integration of the various
orbits. The evaluation of the likelihood for different parameters
\paramsdiff\ of $f$ can be efficiently computed by caching the orbit
integrations of (a) and (b); different source orbits
$(\vecx_p,\vecv_p)$ can also make use of the stored orbits in (a).

\subsection{Action-angle-space model}\label{sec:aamethod}

An alternative description of orbits in a gravitational potential
$\Phi$ is provided by the action-angle formalism (see
\citealt{binneytremaine} for an extensive description). The
Hamiltonian flow in the action-angle formalism is trivial: actions are
conserved and angles increase linearly with time. In
\appendixname~\ref{sec:aa} we summarize a method to calculate
action-angle coordinates from an orbit integration with little extra
work in any (static) potential for any kind of orbit. Therefore, it is
practical to use action-angle coordinates to describe the dynamical
model for a tidal stream in these coordinates and make use of the many
conceptual advantages provided by this formalism.

In the action-angle formalism, a star is removed from the source with
a small offset $(\Delta \vecj,\Delta \veca)(-\Delta t)$ from the
source's actions $\vecj_p$ and angles $\veca_p$ a time $\Delta t$ ago,
similar to the small offset $(\Delta \vecx,\Delta \vecv)(-\Delta t)$
in the real--space model. The offset in the actions corresponds to an
offset in the frequencies $\Delta \veco$; ultimately, this offset is
responsible for the spreading of the stream over long stretches on the
sky. Because both the source's angles and the star's angles increase
linearly with time, albeit at different frequencies, the difference in
the angles also increases linearly in time. Therefore at time $t=0$
the difference in angles is
\begin{equation}
\Delta \veca(t=0) = \Delta \veca(-\Delta t) + \Delta \veco \Delta t\,,
\end{equation}
while the difference in actions is constant
\begin{equation}
\Delta \vecj(t=0) = \Delta \vecj(-\Delta t)\,.
\end{equation}

With these preliminaries, we can write down a generative model for the
positions and velocities ($\vecx_i,\vecv_i$) of stars $i$ in a tidal
stream in the action-angle formalism by specifying: (a) the
gravitational potential $\Phi$, (b) the current action-angle
coordinates of the source $(\vecj_p,\veca_p)(t=0)$, (c) the time
$\Delta t_i$ at which each star was removed from the source, and (d)
the distribution $f(\Delta \vecj,\Delta \veca|\paramsdiff)$,
characterized by some parameters $\paramsdiff$, of initial phase-space
offsets from which the actual offsets $(\Delta \vecj_i,\Delta
\veca_i)$ are drawn. In practice, $f$ can again be taken to be a
Gaussian, although more general models could take the expected
`bowtie' structure of the action distribution into account
\citep{Eyre11a}.

The probability for observing a star today at $(\vecx,\vecv)$ given
items (a) through (d) is then
\begin{align}\label{eq:aadft}
  p(\vecx_i,\vecv_i | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) 
  & = p(\vecj_i,\veca_i | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,,\nonumber\\
  & = p(\vecj_i-\vecj_p,\veca_i-\veca_p-\Delta \veco_i\Delta t_i | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,,\\
  & = f(\Delta \vecj_i,\Delta \veca_i(-\Delta t_i) | \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,,\nonumber
\end{align}
where the first equality holds because the Jacobian of the
$(\vecx,\vecv)\rightarrow(\vecj,\veca)$ transformation is unity and
the second equality is a simple linear transformation. The final
probability can again be evaluated using the parameters
\paramsdiff\ following (d).

As in the real--space formalism, we should integrate
\eqnname~(\ref{eq:aadft}) over the time $\Delta t_i$ at which the star
is removed from the source (cf.~\eqnname~[\ref{eq:realdf}]):
\begin{equation}\label{eq:aadf}
  p(\vecx_i,\vecv_i | \Phi,\vecj_p,\veca_p,\paramsdiff) = \int \dd
  \Delta t_i\, f(\Delta \vecj_i,\Delta \veca_i(-\Delta t_i) |
  \Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff) \,.
\end{equation}
Without further assumptions the action-angle formalism proceeds in the
same way as the real--space formalism, evaluating
\eqnname~(\ref{eq:aadft}) for a grid $j$ of times $\Delta t_i^j$ and
summing $f(\Delta \vecj_i,\Delta \veca_i(-\Delta t^j_i) |
\Phi,\vecj_p,\veca_p,\Delta t_i,\paramsdiff)$ over $j$.

Even without further assumptions about the shape of $f(\Delta
\vecj,\Delta \veca|\paramsdiff)$, the action-angle formalism already
allows for a simplification over the real--space method. Because the
source orbit is specified in action-angle coordinates, it can be
changed \emph{without} expensive orbit integrations of all possible
source orbits. As the source's current phase-space position is a
six-dimensional parameter, this is a major simplification of the
dynamical model when the source orbit is unknown. Note that this
simplification does require us to be able to compute $\veco(\vecj)$,
to calculate the new source frequencies $\veco_p$ when changing the
source actions $\vecj_p$. This mapping can be constructed using a
small number of (action,frequency) calculations covering the range of
possible source actions.

We can simplify \eqnname~(\ref{eq:aadf}) by assuming (a) that
$f(\Delta \vecj,\Delta \veca|\paramsdiff)$ is separable into
$f_{\veca}(\Delta \veca|\paramsdiff)\,f_{\vecj}(\Delta
\vecj|\paramsdiff)$ and (b) that the distribution $f_{\veca}(\Delta
\veca|\paramsdiff)$ is Gaussian. In what follows, we will assume that
this Gaussian is isotropic with dispersion $\sigma_\theta$, but all of
the results are the same after appropriate diagonalization and
rescaling of the angles for more general covariance matrices. In this
case, \eqnname~(\ref{eq:aadf}) can be written as
\begin{equation}
\begin{split}
  p(\vecx_i,\vecv_i & | \Phi,\vecj_p,\veca_p,\paramsdiff) 
   = \\
   & f_{\vecj}(\Delta \vecj_i|\paramsdiff) \int \dd \Delta t_i\,
  \frac{1}{(2\,\pi)^{3/2}\,\sigma_\theta^3}\,\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\left(\Delta \veca_i-\Delta \veco_i\Delta t_i\right)^2\right]\right)\,,
\end{split}
\end{equation}
where the sum in the argument of the exponential is over the three
components of $\Delta \veca_i$ and $\Delta \veco_i$. This integral can
be done analytically, resulting in
\begin{equation}\label{eq:aadfsimple}
\begin{split}
  p(\vecx_i,\vecv_i & | \Phi,\vecj_p,\veca_p,\paramsdiff) 
   = \\
   & f_{\vecj}(\Delta \vecj_i|\paramsdiff) \,  \frac{1}{2\,\pi\,\sigma_\theta^2\,\sqrt{\sum_{R,\phi,Z}\Delta \veco_i^2}}\,\exp\left(-\frac{1}{2\,\sigma_\theta^2}\left[\sum_{R,\phi,Z}\Delta \veca^2_i-\frac{\left(\sum_{R,\phi,Z} \Delta \veco_i\Delta \veca_i\right)^2}{\sum_{R,\phi,Z}\Delta \veco_i^2}\right]\right)\,.
\end{split}
\end{equation}
In this expression, $\Delta X^2 \equiv (X_1-X_2)^2$. The expression in
square brackets is positive by using the Cauchy-Schwarz inequality.

The quantity in the square brackets in \eqnname~(\ref{eq:aadfsimple})
expresses the fact that the present-day difference in angles between a
star in a tidal stream and the source must be linearly related by a
scalar (the time since the removal of the star) to the difference in
frequencies, within a tolerance set by the width of the distribution
of initial angle differences ($\sigma_\theta$). The ratio
$\sum_{R,\phi,Z} \Delta \veco_i\Delta \veca_i/\sum_{R,\phi,Z}\Delta
\veco_i^2$ is the best-guess for $\Delta t_i$ when performing a linear
regression between $\Delta \veca_i$ and $\Delta \veco_i$. The
expression in square brackets is minimized when the correlation
between $\Delta \veca_i$ and $\Delta \veco_i$ is the largest, \ie,
when a single time difference $\Delta t_i$ explains all three of the
angle differences.

The discussion so far has been entirely general and applies to any
sort of stellar debris that originates from a single orbit, without
necessarily generating a one-dimensional structure. In most
observations of tidal streams, however, they appear as one-dimensional
objects rather than the six-dimensional objects that we have
considered so far. The reason for this is both physical and
observational: (a) narrow tidal streams form because a star separates
from the progenitor over time mostly along a single direction in
angle-space: the Hessian matrix of the Hamiltonian, $\partial^2 H /
\partial \vecj \partial \vecj$, has a single large eigenvalue, which
determines the direction along which the stream is disrupted in
angle-space (because $\partial \veco / \partial \vecj$ is equal to the
Hessian); (b) because many tidal streams are dynamically cold with
velocity dispersions $\lesssim10\kms$, observations are unable to
resolve the angle differences along the two eigenvectors of the
Hessian corresponding to the two smallest eigenvalues or the spread in
actions among stream members. If the latter can be resolved, the
stream is four-dimensional. In what follows we will be mainly
concerned with the stream only spreading along a single direction in
angle space.

If the stream is one-dimensional in angle space then the time since
the removal from the progenitor becomes unobservable if the progenitor
orbit is unknown: above, determining the time depended on explaining
the three-dimensional angle difference $\Delta \veca$ with the
three-dimensional frequency difference $\Delta \veco$ using a single
time difference; if two of the angle and frequency differences are
unable to be resolved, then the frequency difference and the time
difference become degenerate (aka, you can draw a straight line
connecting any two points). To make use of this fact to simplify the
dynamical model, we write the frequency difference as
\begin{equation}
  \Delta \veco = \Delta \Omega_1\,\unitvec{e}_1\,,
\end{equation}
where $\unitvec{e}_1$ is a unit vector. Then we write the angle
difference $\Delta \veca$ 
\begin{equation}
  \Delta \veca = \Delta \theta_1\,\unitvec{e}_1 + \Delta \veca_\perp\,\unitvec{e}_1^\perp\,,
\end{equation}
where$\unitvec{e}_1^\perp$ is the orthogonal complement of
$\unitvec{e}_1$ (the two directions perpendicular to $\unitvec{e}_1$)
and $\Delta \veca_\perp$ are the components of $\Delta \veca$ in that
direction. We can then write \eqnname~(\ref{eq:aadfsimple}) as
\begin{equation}\label{eq:aadfsimpler}
\begin{split}
  p(\vecx_i,\vecv_i & | \Phi,\vecj_p,\veca_p,\paramsdiff) 
   = \\
   & f_{\vecj}(\Delta \vecj_i|\paramsdiff) \,  \frac{1}{2\,\pi\,\sigma_\theta^2\,\sqrt{\sum_{R,\phi,Z}\Delta \veco_i^2}}\,\exp\left(-\frac{|\Delta\veca_{i,\perp}|^2}{2\,\sigma_\theta^2}\right)\,.
\end{split}
\end{equation}
\Eqnname~(\ref{eq:aadfsimpler}) is equivalent to
\eqnname~(\ref{eq:aadfsimple}), but it expresses the dynamical model
in a slightly different way. Whereas the angle part of
\eqnname~(\ref{eq:aadfsimple}) allows the correct potential and
progenitor to be inferred by asking whether a single time difference
explains the three angle differences given the three frequencies,
\eqnname~(\ref{eq:aadfsimpler}) `rewinds' the star's orbit along the
direction of $\Delta \veco$ and asks whether the angle difference
perpendicular to $\Delta \veco$ is consistent with the initial angle
spread (because it has not evolved since the removal of the star). In
re-writing this, we have not yet used the approximation that the
stream only forms along a single direction in $\Delta \veco$.

This view of the dynamics of the stream in action-angle coordinates
also shows that the orbit of the source does not have to coincide with
that of the actual progenitor.

The probability in \eqnname~(\ref{eq:aadfsimple}) incorporates and
elucidates commonly used methods for fitting tidal streams. On the one
hand are approaches that search to minimize the spread in energy,
integrals of the motion, or actions (\eg,
\citealt{Binney08a,Penarrubia12a}, Sanderson \etal, 2013, in
preparation). These approaches ignore the correlations between actions
and angles in the stream (the exponential in
\eqnname~(\ref{eq:aadfsimple}), which describes this correlation
through the frequencies) and only use $f_{\vecj}(\Delta
\vecj_i|\paramsdiff)$; finding the gravitational potential that
minimizes the spread in actions optimizes $f_{\vecj}(\Delta
\vecj_i|\paramsdiff)$. Recently, a method was proposed that only uses
the correlation between the actions and the angles, expressed by the
exponential in \eqnname~(\ref{eq:aadfsimple}) \citep{Sanders13b}; this
method does not use the fact that the spread in actions in a tidal
stream is small (expressed by $f_{\vecj}(\Delta
\vecj_i|\paramsdiff)$).

%BOVY: reference point

\acknowledgements It is a pleasure to thank \dots for helpful comments
and assistance. J.B. was supported by NASA through Hubble Fellowship
grant HST-HF-51285.01 from the Space Telescope Science Institute,
which is operated by the Association of Universities for Research in
Astronomy, Incorporated, under NASA contract NAS5-26555. J.B.
acknowledges support from SFB 881 (A3) funded by the German Research
Foundation DFG.

\appendix

\section{An efficient, general method for calculating action-angle coordinates using orbit integration}\label{sec:aa}

The wide-spread use of action--angle coordinates, in many ways the
natural coordinates for studying the orbital structure of galaxies, in
dynamical modeling has been frustrated by the difficulty in
calculating the transformation to and from these coordinates,
$(\vecx,\vecv) \leftrightarrow (\vecj,\veca)$. The most general method
for calculating the transformation $(\vecj,\veca) \rightarrow
(\vecx,\vecv)$ is provided by torus modeling \citep{McGill90a}. In the
context of modeling observational data, computing the transformation
$(\vecx,\vecv) \rightarrow (\vecj,\veca)$ is more important and the
most general method consists of iteratively inverting the torus
modeling \citep{McMillan08a}, which is computationally
expensive. Recently, progress has been made in computing
$(\vecx,\vecv) \rightarrow (\vecj,\veca)$ for Milky-Way-like
potentials by either implicitly or explicitly approximating the
gravitational potential as a St\"{a}ckel potential
\citep{Binney12a,Sanders12a}, which allows action--angle coordinates
to be computed with percent-level errors for moderately eccentric
orbits. However, these methods break down for orbits with radial
and/or vertical actions of a similar magnitude as the angular
momentum---that is, orbits that feel the gravitational potential over
a volume large enough that the St\"{a}ckel approximation fails. This
is especially problematic for the orbits of stars in tidal streams, as
the progenitors of tidal streams are typically on quite eccentric
orbits \citep[\eg,][]{Sanders13a}.

Recently, Fox \& Binney, 2014, in preparation, have suggested a new
method for calculating $\vecj(\vecx,\vecv)$ inspired by torus
modeling. This method requires only a single orbit integration and
allows the actions to be calculated accurately and as precisely as
desired (that is, a convergence criterion can be applied and
convergence can be attained). I describe this method briefly here and
then show how to extend it to also calculate the frequencies and
angles: $(\vecx,\vecv) \rightarrow (\veco,\veca)$.

As in torus modeling, the Fox \& Binney method uses a auxiliary
isochrone potential $\Phi^A$ that should be close to the target
potential $\Phi$ for which the action--angle coordinates should be
calculated (a precise definition of ``close'' will be given
below). Actions and angles $(\vecj^A,\veca^A)$ calculated in the
auxiliary potential can be related to those in the target potential
$(\vecj,\veca)$ by a generating function that can be written as
\citep{McGill90a}
\begin{equation}
  S(\veca^A,\vecj) = \veca^A\cdot\vecj+2\sum_{\vecn > 0} S_{\vecn}(\vecj)\sin(\vecn\cdot\veca^A)\,,
\end{equation}
where the $\vecn > 0$ sum is over the integer two-dimensional
half-plane that excludes the origin (two- rather than
three-dimensional, as the azimuthal angle is perfectly correlated with
the vertical angle in the auxiliary isochrone potential;
\citealt{binneytremaine}), and the $S_{\vecn}$ are functions of the
target actions that are to be determined. The canonical transformation
corresponding to this generating function is
\begin{equation}\label{eq:jjt}
  \vecj^A = \frac{\partial S(\veca^A,\vecj)}{\partial \veca^A} = \vecj + 2\sum_{\vecn > 0} \vecn\,S_{\vecn}(\vecj)\cos(\vecn\cdot\veca^A)\,,
\end{equation}
and
\begin{equation}\label{eq:aat}
  \veca = \frac{\partial S(\veca^A,\vecj)}{\partial \vecj} = \veca^A +2\sum_{\vecn > 0} \frac{\partial S_{\vecn}(\vecj)}{\partial \vecj}\,\sin(\vecn\cdot\veca^A)\,.
\end{equation}

The Fox \& Binney method consists of averaging \eqnname~(\ref{eq:jjt})
over the auxiliary angle along the path of the orbit, such that the
oscillating cosine behavior cancels and the target action is obtained
as the auxiliary-angle-averaged auxilary action. That is,
\begin{equation}\label{eq:jjtint}
  \int \dd \theta_i^A J_i^A = \int \dd \theta_i^A J_i + 2 \sum_{\vecn > 0} \vecn\,S_{\vecn}(J_i)\int \dd \theta_i^A\cos(\vecn\cdot\veca^A)\,,
\end{equation}
where the integral is over auxiliary angles calculated on the path of
the orbit in the target potential. The factor $\vecn\,S_{\vecn}(J_i)$
comes out of the integral as the target actions are conserved along
the orbit. In this Equation, we integrate over the angle coordinate
$\theta_i^A$ associated with the action $J_i$ that we want to compute
(that is, $\theta_R$ for $J_R$ and so on). This is because of the
factor $\vecn$: the condition $\vecn > 0$ means that when calculating
$J_R$ there are terms with $\vecn = (n_R,0)$, which would not cancel
according to the argument below when integrating over $\theta_Z$. For
non-resonant orbits and after a long-enough orbit integration, the
angles can be considered to be independent (that is, the orbit ``fills
its torus''), such that we can write (for example, for $\theta_R$)
\begin{equation}
  \int \dd \theta_R^A\cos(\vecn\cdot\veca^A) = \mathrm{Re}\left[\int \dd \theta_R^A\,\exp(-i\vecn\cdot\veca^A)\right] = \mathrm{Re}\left[\exp(-i\,n_Z\,\theta^A_Z)\int \dd \theta_R^A\,\exp(-i\,n_R\,\theta^A_R)\right]\,.
\end{equation}
As long as the auxiliary angle $\theta_R$ goes through the full range
$0$ to $2\pi$ along the orbit, the oscillatory behavior integrates to
zero and \eqnname~(\ref{eq:jjtint}) simplifies to
\begin{equation}\label{eq:aAIJ}
  J_i = \frac{\int \dd \theta_i^A J_i^A}{\int \dd \theta_i^A}\,,
\end{equation}
that is, the target action is obtained by averaging the auxiliary
action over the auxiliary angle.

The left panel of \figurename~\ref{fig:aAI} shows a running average of
the radial action calculated using \eqnname~(\ref{eq:aAIJ}) along the
orbit of the progenitor in the simulation used in the main body of
this paper. The time in this figure is in units of $35.56\Myr$, so the
orbit is integrated for a total of $7.11\Gyr$. The auxiliary isochrone
potential has a scale parameter of $6.4\kpc$. After a few radial
periods, the radial action quickly converges with a remaining
oscillatory behavior that is primarily a function of the radial angle,
but that becomes smaller over time. With longer integrations and
shorter time-steps, the radial action can converge to any desired
tolerance. The duration of the orbit integration shown in this figure
is what is typically used in this paper and gives a precision of
$\approx2\,\%$ in the actions.

Given the orbit integration already performed for calculating the
actions, we can use \equationname~(\ref{eq:aat}) to calculate the
frequencies and angles corresponding to the point $(\vecx,\vecv)$
(this procedure is similar to that used in \citealt{McMillan08a}). We
first re-write \equationname~(\ref{eq:aat}) in terms of the angle at
time zero and the frequencies
\begin{equation}\label{eq:aat2}
  \veca = \veca(t=0) + \veco(\vecj) \,t = \veca^A +2\sum_{\vecn > 0} \frac{\partial S_{\vecn}(\vecj)}{\partial \vecj}\,\sin(\vecn\cdot\veca^A)\,.
\end{equation}
This a linear system of equations for each point along the orbit with
unknowns $\veca(t=0)$, $\veco(\vecj)$, and the functions
$\frac{\partial S_{\vecn}(\vecj)}{\partial \vecj}$. For a finely
integrated orbit, this system is highly over-constrained (at least
when the number of expansion terms is small) and its solution can be
found by simple linear algebra for any desired number of expansion
terms $\sin(\vecn\cdot\veca^A)$. For extra stability, we integrate the
orbit both forward and backward in time, such that the desired angle
$\veca(t=0)$ is at the center of the dependent variable $t$.

The right panel of \figurename~\ref{fig:aAI} shows the auxiliary
radial angle versus the auxiliary vertical angle along the orbit for
the progenitor used in the main body of this paper. The periodic
behavior in the angles is removed before the linear fit described in
the previous paragraph, which is shown in this figure. It is clear
that the contribution of the oscillatory terms
$\sin(\vecn\cdot\veca^A)$ is small, such that a small number of
expansion terms is sufficient (we limit the terms to $n_R < 4$ and
$n_Z < 4$).

The method described here is completely general, but the only obstacle
to using it is that a sufficiently close isochrone potential has to be
supplied. Sufficiently close means specifically that both the
auxiliary radial and vertical angle have to go through the full range
of $0$ to $2\pi$ along the orbit in the target potential. Whether they
do or not is easy to check and I have found that with limited trial
and error a good enough isochrone potential can be determined quickly
for Milky-Way-like axisymmetric potentials, although it remains to be
checked whether this holds for triaxial potentials as well. For
analyzing stream data, the necessity of this trial-and-error
determinition of the auxiliary potential does not pose a practical
problem, as all of the stream stars are on very similar
orbits. Therefore, the same auxiliary potential can be used for all
action--angle calculations for a given potential and for different
potentials as well, as long as they are not too different. More
widespread use of this technique may require automated methods for
determining a good auxiliary potential. In the unlikely circumstance
that no good auxiliary isochrone potential can be found, the method
discussed here can also be used with other auxiliary potentials for
which the actions and angles can be calculated, such as other
spherical potentials or the family of St\"{a}ckel potentials. This
will, however, add a non-negligible amount of computation for the
necessary numerical integrations.

The action--angle method described in this Appendix is implemented in
the \emph{galpy} Galactic dynamics code\footnote{Available at
  \protect{\url{http://github.com/jobovy/galpy}}~.}.

%BOVY: reference point

\begin{thebibliography}{}

\bibitem[{{Binney} \& {Tremaine}(2008)}]{binneytremaine}
  Binney,~J. \& Tremaine,~S. 2008, Galactic Dynamics: Second Edition
\bibitem[Binney(2008)]{Binney08a}
  Binney,~J. 2008, \mnras, 386, L47
\bibitem[Binney(2012)]{Binney12a}
  Binney,~J. 2012, \mnras, 426, 1324
\bibitem[Carlberg(2009)]{Carlberg09a}
  Carlberg,~R.~G. 2009, \apj, 705, L223
\bibitem[Carlberg(2012)]{Carlberg12a}
Carlberg,~R.~G. 2012, \apj, 748, 20
\bibitem[Dehnen(2000)]{Dehnen00a}
  Dehnen,~W. 2000, \apj, 536, L39
\bibitem[Dehnen(2002)]{Dehnen02a}
  Dehnen,~W. 2002, J.~Comput.~Phys., 179, 27
\bibitem[Eyre \& Binney(2011)]{Eyre11a}
  Eyre,~A. \& Binney,~J. 2011, \mnras, 413, 1852
\bibitem[Helmi \& White(1999)]{Helmi99a}
  Helmi,~A. \& White,~S.~D.~M. 1999, \mnras, 307, 495
\bibitem[Johnston \etal(1999)]{Johnston99a}
  Johnston,~K.~V., Zhao,~H., Spergel,~D.~N., \& Hernquist,~L. 1999, \apj, 512, L109
\bibitem[Koposov \etal(2010)]{Koposov10a}
  Koposov,~S.~E., Rix,~H.-W., \& Hogg,~D.~W. 2010, \apj, 712, 260
\bibitem[McGill \& Binney(1990)]{McGill90a}
  McGill,~C. \& Binney,~J. 1990, \mnras, 244, 634
\bibitem[McMillan \& Binney(2008)]{McMillan08a}
  McMillan,~P.~J. \& Binney,~J.~J. 2008, \mnras, 390, 429
\bibitem[Ngan \& Carlberg(2014)]{Ngan14a}
  Ngan,~W.~H.~W. \& Carlberg,~R.~G. 2014, \apj, submitted
\bibitem[Pe{\~n}arrubia \etal(2012)]{Penarrubia12a}
  Pe{\~n}arrubia, J., Koposov,~S.~E., \& Walker,~M.~G. 2012, \apj, 760, 2
\bibitem[Price-Whelan \& Johnston(2013)]{PriceWhelan13a}
  Price-Whelan,~A.~M. \& Johnston,~K.~V.~2013, \apj, 778, L12
\bibitem[Sanders(2012)]{Sanders12a}
  Sanders,~J. 2012, \mnras, 426, 128
\bibitem[Sanders \& Binney(2013a)]{Sanders13a}
  Sanders,~J. \& Binney,~J. 2013a, \mnras, 433, 1813
\bibitem[Sanders \& Binney(2013b)]{Sanders13b}
  Sanders,~J. \& Binney,~J. 2013b, \mnras, 433, 1826
\bibitem[Teuben(1995)]{Teuben95a}
  Teuben,~P.~J. 1995, in Astronomical Data Analysis Software and Systems IV, 
  ed.~R.~Shaw, H.~E.~Payne and J.~J.~E.~Hayes., PASP Conf Series 77, 398
\bibitem[Tremaine(1999)]{Tremaine99a}
  Tremaine,~S. 1999, \mnras, 307, 877
\bibitem[Yoon \etal(2011)]{Yoon11a}
  Yoon,~J.~H., Johnston,~K.~V., \& Hogg,~D.~W. 2011, \apj, 731, 58
\end{thebibliography}

\begin{figure}[t!]
  \includegraphics[width=0.5\textwidth,clip=]{gd1_evol_xz.ps}
  \caption{Mock tidal stream simulated as described in
    \sectionname~\ref{sec:sim} after $5.011\Gyr$ of evolution in
    Galactocentric $X$ and $Z$ coordinates. The grey dot shows the
    current position of the progenitor cluster and the dashed line
    shows its orbit. The solid line gives the stream track, calculated
    using the method described in \sectionname~\ref{sec:track}. The
    inset shows a zoomed-in version of part of the leading stream that
    more clearly shows the offset between the stream track and the
    progenitor orbit.}\label{fig:gd1_xz}
%python plot_stream.py ../tex/gd1_evol_xz.ps
\end{figure}

\begin{figure}[t!]
  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aajrjz.ps}
  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaoroz.ps}
  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaraz.ps}\\
  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aajrjp.ps}
  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaorop.ps}
  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaarap.ps}
  \caption{Simulated tidal stream from \sectionname~\ref{sec:sim}
    after $5.011\Gyr$ of evolution in action--frequency--angle
    coordinates. The top row displays projections of the actions
    (left), frequencies (middle), and angles (right) in the $R$ and
    $Z$ directions; the bottom panels shows the same for $R$ and
    $\phi$. While the action distribution has a quite complicated
    form, the frequency distribution is close to one-dimensional and
    the leading and trailing parts of the stream clearly separate. In
    the right panels, the angles of the progenitor have each been
    shifted to $\pi$ and the dashed lines show the direction of the
    progenitor's frequency vector. The misalignment between the
    progenitor's frequency vector and the mean frequency vector of the
    debris is $0.18^\circ$.}\label{fig:gd1_jt}
%python plot_stream.py ../tex/gd1_evol_aajrjz.ps
%python plot_stream.py ../tex/gd1_evol_aaoroz.ps
%python plot_stream.py ../tex/gd1_evol_aaaraz.ps
%python plot_stream.py ../tex/gd1_evol_aajrjp.ps
%python plot_stream.py ../tex/gd1_evol_aaorop.ps
%python plot_stream.py ../tex/gd1_evol_aaarap.ps
\end{figure}

\begin{figure}[t!]
  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timeshist.ps}
  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timesdO.ps}
  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_timesda.ps}
  \caption{Properties of the simulated tidal stream of
    \sectionname~\ref{sec:sim} as a function of time. Stripping times
    are calculated by linear regression of the angle difference
    between a stream particle and the progenitor versus the equivalent
    frequency difference. The distribution of stripping times is shown
    in the left panel.  While the stripping process happens in bursts
    with a period of roughly $400\Myr$ (the radial period of the
    progenitor), it can be approximated as constant (more
    coarsely-binned solid histogram). The magnitude of the frequency
    offset between stream particles and the progenitor is shown in the
    middle panel. While particles stripped at pericenter have larger
    frequency offsets than those between pericenter passages, there is
    no secular trend and the typical frequency offset (averaged over a
    radial period) can be considered constant with time. The right
    panel shows the magnitude of the angle offset between stream
    particles and the progenitor as a function of stripping
    time. There is a spread in the angle offset at all stripping times
    because of the dispersion in frequency offsets, but typically
    stars that have been stripped earlier reach larger angle
    offsets. That is, self-sorting happens only to a small
    degree.}\label{fig:gd1_times}
%python plot_stream.py ../tex/gd1_evol_timeshist.ps
%python plot_stream.py ../tex/gd1_evol_timesdO.ps
%python plot_stream.py ../tex/gd1_evol_timesda.ps
\end{figure}

\begin{figure}[t!]
  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aadohist.ps}
  \caption{Distribution of the magnitude of frequency offsets between
    stream particles and the progenitor along the direction of the
    mean frequency offset. The dashed line is a Gaussian fit to this
    distribution, while the solid line is a fit of our preferred form
    for this distribution, given in \equationname~(\ref{eq:do}). Both
    forms provide adequate fits to the
    distribution. }\label{fig:gd1_dohist}
%python plot_stream.py ../tex/gd1_evol_aadohist.ps
\end{figure}

\begin{figure}[t!]
  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparopar.ps}
  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparoperp.ps}\\
  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaaparaperp.ps}
  \includegraphics[width=0.32\textwidth,clip=]{gd1_evol_aaapartime.ps}
  \caption{Stream properties as a function of angle along the
    stream. The top, left panel shows the magnitude of frequency
    offsets along the stream. The gray line shows the mean predicted
    frequency offset calculated using the framework of
    \sectionname\sectionname~\ref{sec:modeloa} and \ref{sec:track},
    while the grayscale bands show the $68\,\%$ and $95\,\%$ limits of
    the predicted distribution. The top, right panel shows the
    magnitude of frequency offsets perpendicular to those in the left
    panel. The predicted distribution is constant as a function of
    angle and therefore not shown. The bottom, left panel shows the
    angle offset perpendicular to the stream and the grayscale bands
    show $68\,\%$ and $95\,\%$ limits of the predicted distribution of
    angle offsets; the white dashed line shows the $68\,\%$ limit
    calculated using a simplified method. The bottom, right panel
    shows the stripping time. The gray line and grayscale bands again
    show the mean and $68\,\%$ and $95\,\%$ confidence limits of the
    predicted distribution. While the ``burstiness'' related to the
    energetic stripping at pericenter that is apparent in the
    simulated data's distributions is (by design) absent in the model,
    the overall scale and the mean trends with angle are well
    represented by the model.}\label{fig:gd1_apar}
%python plot_stream.py ../tex/gd1_evol_aaaparopar.ps
%python plot_stream.py ../tex/gd1_evol_aaaparoperp.ps
%python plot_stream.py ../tex/gd1_evol_aaaparaperp.ps
%python plot_stream.py ../tex/gd1_evol_aaapartime.ps
\end{figure}

\begin{figure}[t!]
  \includegraphics[width=\textwidth,clip=]{gd1_evol_lb.ps}
  \includegraphics[width=\textwidth,clip=]{gd1_evol_ld.ps}
  \caption{Simulated stream after $5.011\Gyr$ as a function of Galactic
    longitude. The top panel shows the stream in Galactic latitude and
    the bottom panel shows the distance from the Sun. The progenitor
    is shown as the large gray dot and its orbit as the dashed
    lines. The predicted track of the stream is shown as the solid
    line. Insets show various zoomed-in parts of the stream to show
    the stream, progenitor orbit, and predicted stream track up
    close.}\label{fig:gd1_lbd}
%python plot_stream.py ../tex/gd1_evol_lb.ps
%python plot_stream.py ../tex/gd1_evol_ld.ps
\end{figure}
\begin{figure}[t!]
  \includegraphics[width=\textwidth,clip=]{gd1_evol_lvlos.ps}
  \includegraphics[width=\textwidth,clip=]{gd1_evol_lpmll.ps}
  \includegraphics[width=\textwidth,clip=]{gd1_evol_lpmbb.ps}
  \caption{Simulated stream velocities as a function of Galactic
    longitude. Same as \figurename~\ref{fig:gd1_lbd}, but for the
    stream in heliocentric line-of-sight velocity $V_{\mathrm{los}}$
    (top panel), proper motion in the direction of Galactic longitude
    (middle panel), and in the direction of Galactic latitude (bottom
    panel). Insets again show various interesting parts of the stream
    up close.}\label{fig:gd1_lbv}
%python plot_stream.py ../tex/gd1_evol_lvlos.ps
%python plot_stream.py ../tex/gd1_evol_lpmll.ps
%python plot_stream.py ../tex/gd1_evol_lpmbb.ps
\end{figure}

\begin{figure}[t!]
 \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_2Gya_xz.ps}
 \includegraphics[width=0.48\textwidth,clip=]{gd1_evol_2Gya_xz_sim.ps}
  \caption{Mock stream data generated from the model. The left panel
    shows the same $N$-body stream as shown in previous figures, but
    $1.96\Gyr$ earlier in its evolution. Symbols, lines, and the
    stream model are the same as those in
    \figurename~\ref{fig:gd1_xz}, except that the model progenitor
    phase-space location from \figurename~\ref{fig:gd1_xz} has been
    integrated backward in time for $1.96\Gyr$ (no other changes were
    made to the model's parameters). The right panel shows mock stream
    data drawn from the model using the procedures discussed in
    \sectionname~\ref{sec:mock}. Overall, the spatial distribution of
    the mock data is very similar to that of the $N$-body
    stream.}\label{fig:gd1_sim}
%python plot_stream_2Gya.py ../tex/gd1_evol_2Gya_xz.ps
%python plot_stream_2Gya.py ../tex/gd1_evol_2Gya_xz_sim.ps
\end{figure}

\begin{figure}[t!]
 \includegraphics[width=0.48\textwidth,clip=]{aAI_jr.ps}
 \includegraphics[width=0.48\textwidth,clip=]{aAI_araz.ps}
  \caption{Action--angle calculations using an orbit-integration-based
    method. The right panel shows the cumulative mean of the radial
    actions calculated in an auxiliary isochrone potential calculated
    along the orbit of the progenitor used in the main body of this
    paper. The color-coding corresponds to the radial angle
    (calculated in the isochrone potential) along the orbit. The time
    is in units of $35.56\Myr$ and the radial action is normalized by
    its value at $t=200$. The left panel shows the radial and vertical
    angles calculated in the auxiliary isochrone potential, with
    color-coding corresponding to the azimuthal angle. The wiggles in
    the angles are due to the isochrone potential not being the
    correct potential, but these wiggles are fit and removed to
    calculate the true angles.}\label{fig:aAI}
%python plot_aAI.py ../tex/aAI_jr.ps
%python plot_aAI.py ../tex/aAI_araz.ps
\end{figure}

\end{document}